\section{An Executable Trust Specification}
\label{sec:bdd}

The anti-hallucination guarantee is not a slogan; it is encoded as
behavior-driven tests in Gherkin~\cite{gherkin}, written \emph{in French} so
that a jurist---not only an engineer---can read, challenge, and sign off on
them. The suite runs on every change, and \emph{a single fabricated article
turns the build red}. Turning the invariant into a red/green signal makes it
continuously checkable rather than merely asserted, and hands a non-technical
domain expert a direct grip on the system's behaviour.

Three scenarios cover the invariant. (i)~A verified citation must be presented
\emph{with} its source. (ii)~A question that elicits a fabricated
article---the ``Christmas bonus'', for which the model invents
\emph{article~L.\,4321-7}---must be refused, and no article absent from the
sources may be shown as verified. (iii)~A question with no verifiable source
must be refused rather than answered from parametric memory.

\begin{lstlisting}[style=lr]
Scenario: No invented citation
  Given a citizen question "Ai-je droit a la prime de Noel ?"
  When the system answers
  Then the answer must be an explicit refusal
  And no article absent from the sources may be
      presented as verified
\end{lstlisting}

Because the offline mock verifier (\S\ref{sec:verify}) exercises the same
decision logic as the live backend, these scenarios are deterministic: they
pin the \emph{behaviour} of the trust layer independently of which model or
accelerator sits behind it, so what turns the suite red is a regression in the
pipeline---not merely a model that happens to answer well today.

The other two scenarios pin the positive and the no-source cases:

\begin{lstlisting}[style=lr]
Scenario: A verified citation is always sourced
  Given a question with a known in-force article
  When the system answers
  Then every article shown as verified
       resolves in the database
  And each carries a clickable Legifrance link

Scenario: No source, no answer
  Given a question with no verifiable legal source
  When the system answers
  Then the answer must be an explicit refusal
\end{lstlisting}

\subsection{The harness}
The scenarios run under \texttt{behave}: French \emph{Soit\,/\,Quand\,/\,Alors}
steps drive the pipeline in offline mock mode, \texttt{environment.py} wires
the deterministic verifier, and the suite is a continuous-integration gate.
Because the steps read as domain language rather than code, a jurist can
\emph{author} new obligations---each becoming a test the system must satisfy on
every commit.
