\section{Evaluation on 100 Citizen Questions}
\label{sec:eval}

\evalModeBanner

We evaluate the fail-closed contract on a suite of \evalN real French
citizen legal questions, stratified into three buckets: \emph{answerable}
(\evalAnswerableN questions for which a well-known in-force code article
applies), \emph{trap} (\evalTrapN questions phrased to elicit a
plausible-but-nonexistent right or article), and \emph{open} (\evalOpenN broad
or multi-source questions). Every question is run through the \emph{identical}
pipeline (backend \evalBackend, model \evalModel; verifier \evalVerifier). All
figures below are generated directly from that run
(\texttt{tools/bench\_eval.py} $\to$ \texttt{metrics.tex}); the
database is the sole source of truth for whether a cited article exists and is
in force.

\paragraph{The invariant holds end-to-end.}
Across \evalGraded graded questions, \evalLeak citations were surfaced to the
user as \emph{verified} while failing to resolve to an in-force article. In
other words, the ``no verifiable source, no answer'' contract was never
violated: of the \evalCitTotal citations the model emitted, the \evalCitVerified
that resolved in the database were the only ones ever presented as sourced, and
each of the \evalCitFailed unresolved citations forced a refusal instead of a
fabrication.

\paragraph{The guard fires in the wild.}
\evalGuardFired questions were refused \emph{because} at least one cited article
failed verification, catching \evalDistinctInvented distinct non-resolvable
article numbers---exactly the confident fabrications a fluent model produces.
Overall the system refused \evalRefusalRate of questions. On the trap bucket it
correctly refused \evalTrapRefused/\evalTrapN (\evalTrapRate); on the answerable
bucket it returned a database-verified answer for
\evalAnswerableAnswered/\evalAnswerableN (\evalAnswerableRate).

\paragraph{Refusal is bounded by coverage, not only fluency.}
The answerable-bucket gap is a \emph{systems} result as much as a model one: a
refusal also fires when the model cites the correct rule in a form the verifier
cannot resolve (a numbering variant, or an article outside the mirrored
sub-corpus). Answer rate is therefore lower-bounded by corpus coverage and
citation normalisation, not by generation quality alone---the direction of
future work in Section~\ref{sec:limits}. Per-citation checks ran at interactive
latency (median \evalLatMedian\,s, p95 \evalLatPToNine\,s end-to-end;
\evalTokps~tok/s) over the 253-table corpus without full-table scans
(Section~\ref{sec:verify}).

\begin{table}[t]
\centering
\caption{Le Rapporteur on \evalN citizen questions (mode \evalMode, backend
\evalBackend, \evalDate). Generated from the run; the fail-closed contract
surfaced \evalLeak fabricated citation as verified.}
\label{tab:eval}
\small
\begin{tabular}{@{}lr@{}}
\toprule
Metric & Value \\
\midrule
Questions ($n$) & \evalN \\
\quad answerable / trap / open & \evalAnswerableN\,/\,\evalTrapN\,/\,\evalOpenN \\
\midrule
Fabrications surfaced as verified & \textbf{\evalLeak} \\
Refusals triggered by a bad citation & \evalGuardFired \\
Distinct fabricated articles caught & \evalDistinctInvented \\
Citations verified / total & \evalCitVerified\,/\,\evalCitTotal \\
\midrule
Overall refusal rate & \evalRefusalRate \\
Trap correctly refused & \evalTrapRefused/\evalTrapN\ (\evalTrapRate) \\
Answerable database-verified & \evalAnswerableAnswered/\evalAnswerableN\ (\evalAnswerableRate) \\
\midrule
Latency median / p95 & \evalLatMedian\,s\,/\,\evalLatPToNine\,s \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Two representative cases.}
Beyond the aggregates, both the power \emph{and} the boundary of existence
checking appear directly. When the model invents a non-existent
authority---the ``Christmas bonus'' $\to$ \emph{article L.\,4321-7} (type~E1,
\S\ref{sec:taxonomy})---the verifier resolves it to \texttt{exists=false} and
the system refuses rather than fabricate: precisely the error class a fluent
model produces and a database check catches. When instead, asked for the legal
working week, the model answers \emph{article L.\,3132-2}---which exists and is
in force, but governs weekly \emph{rest}---rather than the correct
\emph{L.\,3121-27}, the citation is a type-E4 error: real, in force, yet
off-topic. An existence check alone does \emph{not} catch it, and we do not
claim it does; it marks exactly the boundary between the guarantee we provide
today and the semantic guarantee of Section~\ref{sec:limits}.
