\section{Limitations and Future Work}
\label{sec:limits}

\paragraph{Existence is not semantics.}
Our verifier checks that a cited article \emph{exists} and is \emph{in force};
it does not yet check that the article's \emph{content} supports the answer's
claim. As \S\ref{sec:eval} shows, a real but off-topic citation (type~E4)
passes an existence check. Closing this gap is a natural-language inference
problem: treat the article text as the premise and the answer's claim as the
hypothesis, and require \emph{entailment} before releasing the
citation~\cite{williams2018mnli}. Legal-domain encoders such as
LEGAL-BERT~\cite{chalkidis2020legalbert} and fact-verification pipelines in the
FEVER tradition~\cite{thorne2018fever} are the obvious building blocks, and
benchmarks like LegalBench~\cite{guha2023legalbench} would let us measure
progress. This would upgrade the guarantee from ``the source exists'' to ``the
source says so''.

\paragraph{Grounding: generate-then-verify $\rightarrow$ retrieve-then-generate.}
The current system \emph{generates then verifies}; it does not yet retrieve
before generating. The planned grounding is deliberately \emph{lexical} rather
than vectorial: a full-text search---via a metadata-aware SQL tool or
PostgreSQL \texttt{to\_tsvector}, ranked with BM25~\cite{robertson2009bm25}---
returning top-$k$ passages that condition generation, complementing dense
retrieval~\cite{karpukhin2020dpr, lewis2020rag}. The obstacle is not a vector
store but an efficient full-text \emph{index} over a very large, versioned
article table; indexing the entire corpus is the real wall, so we would scope
to a sub-corpus (e.g., a single code) first.

\paragraph{Model and hardware.}
Mistral is not served by our live Qualcomm endpoint (Llama-3.1-8B only); a
Mistral deployment through the portable runtime on AMD is future work, as is a
kernel-level (Mojo) hot-path study to support a \emph{measured} perf/watt
claim---which we have deliberately avoided asserting without data
(\S\ref{sec:sovereign}).

\paragraph{Verifier soundness and coverage.}
Our guarantee is only as good as the verifier's soundness and the corpus's
completeness. The extractor targets the canonical \emph{``article~X du
Code~Y''} form and will miss unusual citation phrasings---which, being
unextracted, are simply not vouched for, a fail-closed omission rather than a
false endorsement. Coverage beyond codified statute (case law, EU law,
circulars) and robustness to multilingual input remain open.

\paragraph{From refusal to repair.}
A refusal is safe but blunt. A natural next step is \emph{repair}: when a draft
cites a non-existent article, use the grounding index above to retrieve the
provision the citizen most likely needs and regenerate, moving the system from
``I cannot confirm this'' toward ``here is the text that actually
applies''---without ever relaxing the release predicate. Refusal remains the
safe default; repair is a strict improvement only if the retrieved,
re-generated citation itself passes verification.
