\section{Citation Verification over a Legal Database}
\label{sec:verify}

The verifier answers a deceptively simple query: \emph{does article
$\langle$num$\rangle$ of $\langle$code$\rangle$ exist and is it in force?} Doing
this at interactive latency over Canutes required care.

\paragraph{Index-friendly citation matching.}
Légifrance encodes article numbers inconsistently (e.g., \texttt{L3121-27},
\texttt{L.\,3121-27}, \texttt{L. 3121-27}). A naive predicate
(\texttt{num \textasciitilde{} '.*3121-27.*'}) forces a sequential scan and a
regular-expression match over the whole \texttt{legifrance.article} table,
which in our environment hit a statement timeout. Instead, we expand the
citation into a small, closed set of \emph{exact} string variants and match
with equality against the (indexable) \texttt{num} column:

\begin{lstlisting}[style=lr, language=SQL]
-- variants = {"L. 3121-27", "L.3121-27", "L3121-27", ...}
SELECT id, num, data
FROM   legifrance.article
WHERE  num = ANY(%s)          -- equality, index-friendly
  AND  data::text ILIKE %s    -- code name appears in the article JSON
LIMIT  200;
\end{lstlisting}

\paragraph{Code disambiguation and accents.}
The code name (e.g., \emph{Code du travail}) is matched by \texttt{ILIKE}
against the article's JSON payload, then re-checked in Python with an
accent-insensitive normalization (Postgres \texttt{ILIKE} is not
accent-folding). This two-stage filter keeps the SQL cheap while remaining
correct on inputs such as \emph{Code de la s\'ecurit\'e sociale}.

\paragraph{State-aware selection.}
Among the surviving rows we prefer the \texttt{VIGUEUR} (in-force) version;
if present, we return its canonical \texttt{LEGIARTI} identifier, from which we
build a real, clickable Légifrance URL, plus a short text excerpt. If no row
matches the code, the verifier returns \texttt{exists=false}, which the
pipeline treats as a refusal trigger. Crucially, \emph{every failure mode maps
to ``not verified''}: this is what makes the layer fail-closed.

\paragraph{Offline determinism.}
For demos and tests without network access, a mock verifier answers from a
small local set of articles known to be in force. The same code path yields a
deterministic, reproducible refusal for a fabricated article---essential for
the executable specification of \S\ref{sec:bdd}.
